Declarative runtime reconfiguration of pulverised distributed systems \thanks{Identify applicable funding agency here. If none, delete this.}

### 2023-05-11 paper-2023-acsos-pulverization.tex
RW: more on app description and pulverisation

This work provides 
 \acp{dsl}
 for specifying 
 distributed systems and
 corresponding deployment plans,
 and a framework for executing
 deployment plans and reconfigurations.
%
Therefore, this work deals 
 with multiple problems and perspectives,
 investigated by multiple research threads.
%
In~\cite{DBLP:journals/jss/ArcangeliBL15},
the problem of (automatic) deployment of distributed software system is studied by considering
 (i) the nature of software to be deployed,
 (ii) the nature of the deployment domain (i.e. the available infrastructure),
 (iii) how the deployment is designed,
 and 
 (iv) how the deployment is performed.
%
Accordingly, 
 we cover related works about
 languages for 
 specifying applications (\Cref{sec:rw:whatsw}),
 languages for specifying infrastructures and deployments (\Cref{sec:rw:depdesc}),
 and
 techniques for automatic deployment and reconfiguration (\Cref{sec:rw:autodep}).

\subsection{Application description languages: component models and pulverisation}
\label{sec:rw:whatsw}
%
A comprehensive survey on component-based software engineering is provided in~\cite{vale2016component-based-se}.

%
The approach of specifying in a single codebase 
 parts of the structure, behaviour, or interaction 
 for a whole distributed system
 is also related to \emph{macroprogramming}~\cite{Casadei2023macro}
 and, in particular, to \emph{multi-tier programming} paradigm~\cite{DBLP:journals/csur/WeisenburgerWS20},
 where the deployment units for different tiers (e.g., client and server tiers; or view, business logic, and data tiers) are obrained by compilation or interpretation of a single codebase.

@@

\subsection{Automatic deployment and reconfiguration}
\label{sec:rw:autodep}
%
A comprehensive conceptual framework
 on automatic deployment of distributed systems
 is provided in a 2015 survey by Arcangeli et al. \cite{DBLP:journals/jss/ArcangeliBL15}.
%
A more recent survey by Coullon et al. focusses
 on formal techniques for verifying the correctness of reconfigurations
 in component-based distributed software systems~\cite{coullon2023swreconfig}.
%
In the \emph{osmotic computing} approach~\cite{DBLP:journals/computer/VillariFDRJR19},
 microservices (the solvent)
 can migrate across 
 the edge-cloud infrastructure (the solution),
 passing through layer boundaries (the semi-permeable membranes),
 in order to keep a balance in the desired properties (the solute).

>>>>

This work provides 
 \acp{dsl}
 for specifying 
 distributed systems and
 corresponding deployment plans,
 and a framework for executing
 deployment plans and reconfigurations.
%
Therefore, we provide an original solution
 to different issues.
%
In~\cite{DBLP:journals/jss/ArcangeliBL15},
the problem of (automatic) deployment of distributed software system is studied by considering
 (i) the nature of software to be deployed,
 (ii) the nature of the deployment domain (i.e. the available infrastructure),
 (iii) how the deployment is designed,
 and 
 (iv) how the deployment is performed.
%
Accordingly, 
 we cover related works about
 languages for 
 specifying distributed systems (\Cref{sec:rw:whatsw}),
 languages for specifying infrastructures and deployments (\Cref{sec:rw:depdesc}),
 and
 techniques for automatic deployment and reconfiguration (\Cref{sec:rw:autodep}).

\subsection{Application description languages: component-based software engineering and the pulverisation model}
\label{sec:rw:whatsw}
%
%% A first distinction revolves around whether 
%% the distributed software to be deployed 
%% is a monolith 
%% or is already broken into different components.
%
We consider a distributed application 
 as a graph of deployable units.
%
This partitioning may be 
 manually specified at development time
 (e.g., by explicitly defining and packaging different components)
 or automatically defined
 through \emph{application partitioning} approaches~\cite{DBLP:journals/jnca/LiuASGBQ15}. %exploiting the natural modularity of code (depending on the paradigm and language adopted).

In our approach, the application partitioning into components is manually specified through a \ac{dsl}.
%
Therefore, it can be framed in the context of \emph{component-based software engineering},
 with ~\cite{vale2016component-based-se} providing a comprehensive review.
%
The specification or design of distributed systems
 may leverage
 component models~\cite{DBLP:journals/tse/CrnkovicSVC11},
 architectural description languages~\cite{DBLP:journals/tse/MedvidovicT00},
 or frameworks.
%

Specifically,
 this work generalises over the \emph{pulverisation model}~\cite{FI2020-pulverization,IEEE-IoTJ-pulverization-simulation},
 whereby a large-scale cyber-physical system
 is partitioned into a graph of devices
 where each device is split into five deployable components:
 (i) sensors interface, 
 (ii) actuators interface, 
 (iii) behaviour, 
 (iv) state, and 
 (v) communication component.
%
In~\cite{IEEE-IoTJ-pulverization-simulation},
 a methodology on top of the pulverisation model
 is proposed
 for generating and testing deployments through simulation.

Also,
 the approach of specifying in a single codebase 
 parts of the structure, behaviour, and/or interaction 
 for a whole distributed system
 is also related to \emph{macroprogramming}~\cite{Casadei2023macro}
 and, in particular, to \emph{multi-tier programming} paradigm~\cite{DBLP:journals/csur/WeisenburgerWS20},
 where the deployment units for different tiers (e.g., client and server tiers; or view, business logic, and data tiers) are obrained by compilation or interpretation of a single codebase.

@@

\subsection{Automatic deployment and reconfiguration}
\label{sec:rw:autodep}
%
A comprehensive conceptual framework
 on automatic deployment of distributed systems
 is provided in a 2015 survey by Arcangeli et al. \cite{DBLP:journals/jss/ArcangeliBL15}.
%
A more recent survey by Coullon et al. focusses
 on formal techniques for verifying the correctness of reconfigurations
 in component-based distributed software systems~\cite{coullon2023swreconfig}.
%
\meta{TODO: say something about our positioning wrt verification of reconfigurations}

In the \emph{osmotic computing} approach~\cite{DBLP:journals/computer/VillariFDRJR19},
 microservices (the solvent)
 can migrate across 
 the edge-cloud infrastructure (the solution),
 passing through layer boundaries (the semi-permeable membranes),
 in order to keep a balance in the desired properties (the solute).
